Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-17 paper.tex
Mention PVS also in intro

In many proof assistants, simple \enquote*{white-box} automation tactics based on tableau-style tree search play a substantial role in practice: Isabelle's (TODO ref) and Rocq's (TODO ref) auto, ACL2's (TODO ref) waterfall, Lean's (TODO ref) Aesop (TODO ref), etc.
These tactics operate on \emph{goals} $Γ ⊢ A$, where the \emph{context} $Γ$ is a list of assumptions and $A$ is the proposition to be proved.
Faced with an initial goal, the tactics try various user-specified \emph{rules}, which correspond to admissible inferences in the prover's logic.
A successful rule reduces the current goal to zero or more subgoals, which are solved recursively until either a full proof is found or no rule applies to any open goal or the search times out.

@@

ACL2 and Aesop additionally provide special support for \emph{forward rules}.
Given, for example, the transitivity lemma $\All{x~y~z}{x < y → y < z → x < z}$, a forward rule run on the goal $a < b,\, b < c ⊢ A$ produces the single subgoal $a < b,\, b < c,\, a < c ⊢ A$.
Hence forward rules can be used to establish additional facts that may be relevant for other rules.
For example, the Aeneas project (TODO ref) uses Aesop's forward rules \jcom{(ensure that this is accurate)} in its proof automation to establish facts that are later used by an arithmetic decision procedure.

However, Aesop's (and ACL2's) support for forward rules was, until recently, quite basic.
In particular, each forward rule was tried on each goal separately (subject to pre-filtering by a discrimination tree (TODO ref) index).
This is wasteful since in practice, many rules produce subgoals with a similar---or even the same---context as the initial goal.
For example, the ∧-introduction rule affects only the goal's target, not its context.
Still, all forward rules applied to the initial $Γ ⊢ A ∧ B$ would be tried again on the subgoals $Γ ⊢ A$ and $Γ ⊢ B$, with predictably little effect.

>>>>

In many proof assistants, simple \enquote*{white-box} automation tactics based on tableau-style tree search play a substantial role in practice: Isabelle's (TODO ref) and Rocq's (TODO ref) auto, ACL2's (TODO ref) waterfall, various PVS (TODO ref) tactics, Lean's (TODO ref) Aesop (TODO ref), etc.
These tactics operate on \emph{goals} $Γ ⊢ A$, where the \emph{context} $Γ$ is a list of assumptions and $A$ is the proposition to be proved.
Faced with an initial goal, the tactics try various user-specified \emph{rules}, which correspond to admissible inferences in the prover's logic.
A successful rule reduces the current goal to zero or more subgoals, which are solved recursively until either a full proof is found or no rule applies to any open goal or the search times out.

@@

ACL2, PVS and Aesop provide special support for \emph{forward rules}.
Given, for example, the transitivity lemma $\All{x~y~z}{x < y → y < z → x < z}$, a forward rule run on the goal $h₁ : a < b,\, h₂ : b < c ⊢ A$ produces the single subgoal $h₁ : a < b,\, h₂ : b < c,\, h₃ : a < c ⊢ A$.
Hence forward rules can be used to establish additional facts that may be relevant for other rules.
For example, the Aeneas project (TODO ref) uses Aesop's forward rules \jcom{(ensure that this is accurate)} in its proof automation to establish facts that are later used by an arithmetic decision procedure.

However, Aesop's support for forward rules was, until recently, quite basic.
In particular, each forward rule was tried on each goal separately (subject to pre-filtering by a discrimination tree (TODO ref) index).
This is wasteful since in practice, many rules produce subgoals with a similar---or even the same---context as the initial goal.
For example, the ∧-introduction rule affects only the goal's target, not its context.
Still, all forward rules applied to the initial $Γ ⊢ A ∧ B$ would be tried again on the subgoals $Γ ⊢ A$ and $Γ ⊢ B$, with predictably little effect.
